---
title: "EDA"
author: "Phillip Wittel"
date: "4/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
masterTable <- read_csv("Data Clean/master_table.csv")

MT_gdpPop <- masterTable %>% 
  mutate("Per capita GDP (in USD 2015)" = masterTable$`GPD (in USD 2015)`/ Population_2010)

MT_gdpPop %>% filter(is.na(Population_2010) == 0) %>% 
  pull(Entity) %>% 
  str_detect("Island") %>% 
  sum()

MT_gdpPop %>% filter(is.na(Population_2010) == 1) %>% 
  pull(Entity) %>% 
  str_detect("Island") %>% 
  sum()

```
There are 12 Enities that have missing values in population with the word "Island", compared to 2 Entities that do to have values for population with the word "island"

```{r}

MT_gdpPop %>% filter(MT_gdpPop %>% pull(Entity) %>% str_detect("Congo"))

MT_gdpPop %>% filter(MT_gdpPop %>% pull(Entity) %>% str_ends("union"))

```
We have duplicates of the Democratic Republic of the Congo, and also a third. I am okay with remove this group of data points as a.) there was a civil war in DRC and b.) we have a lot of missing data points for these nations; a lot of imputation is going to be required for these nations. 
We also have a duplicate of Reunion, since it uses an acsent on one. 


```{r}

plot(y = MT_gdpPop %>% pull(`Plastic waste generation (tonnes, total)`) %>% log10(), 
     x = MT_gdpPop %>% pull(Population_2010) %>% log10)

plot(y = MT_gdpPop %>% pull(`Plastic waste generation (tonnes, total)`) %>% log10(), 
     x = MT_gdpPop %>% pull(`GPD (in USD 2015)`) %>% log10)

plot(y = MT_gdpPop %>% pull(`Plastic waste generation (tonnes, total)`) %>% log10(), 
     x = MT_gdpPop %>% pull(`Per capita GDP (in USD 2015)`) %>% log10)

plot(y = MT_gdpPop %>% pull(`Per capita plastic waste (kg/person/day)`) %>% log10(), 
     x = MT_gdpPop %>% pull(`Per capita GDP (in USD 2015)`) %>% log10())

```
