---
title: "Statisticial Analysis of the Factors that Correlate to Plastic Waste Generation"
author: "P. Wittel"
date: "4/13/2022"
bibliography: references.yaml
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

------------------------------------------------------------------------

## Abstract

|       In this study, we investigate possible socioeconomic factors that influence plastic waste generated by political entity, using data available in 2010. Imputation was used to increase the available dataset. showing that...

## Introduction

|       Plastic Waste and plastic pollution is a growing global problem. Current end-of-life processing of plastic are insufficient for dealing with the mass of plastic production in a environmentally sustainable manner. Only a total of XXX % of Plastic ever generated has been recycled since 1955, and today's recycling rate is at XXX %. Incineration is another end-of-life process for plastics, however incinerated plastic contaminate the surrounding air quality, effecting the health of local residents and contributing green house gases. The most common end-of-life process for plastic is disposal, which includes both managed and un-managed disposal. However, most commercial plastics are not naturally degradable and plastic waste that is exposed to UV radiation will break down to micro-plastics which has a negative effect on both environmental health as well as human health. These issues of waste management are compound by the annual increase in the production of plastics each year. Plastic production has increased from XXX in 1950 to XXX in 2021, with little evidence of slowing down.

|       Methods to understand the Plastic waste problem through statistical means have focused on the effect of waste management policies, Plastic waste damage to the environment, or measuring the quantity of plastic that is leaked into the environment. However there are few studies that focus on the big-picture socioeconomic effects of plastic waste generation, such as population, GPD, and quality of life. These factors would presumably would generate demand and thus a market for plastic products, yet this relationship is has not been show statistically.

|       To address the issue of plastic waste, this study aims to investigate the the historic economical and social factors of political entities in 2010 that contributed to the generation of plastic waste in 2010. Entities in this case are recognized semi-independent political structures; examples include nation states, city-states, and overseas territories.

## Methodology

|       Data was pull from open-sourced data bases that were freely available online. However this limited the scope of this investigation. We pulled plastic generation and plastic waste from our *Our World in Data* and economic and social metrics such as GDP, population, economic growth, and wealth disparage from *the World Bank*. We explored several machine-learning methods, including supervised and un-supervised methods, using the imputed data and investigated different metrics to compare the models created.

|       The purpose of this model is explanation, therefore we do not need to use a typical 70:30 split of the data for a training dataset and a testing dataset. However to ensure that the models made are not prone to over-fitting the available data, we used the most through cross-validation method available to us, LOOCV. This ensures that the final model does not overfit to high-impact data points. We were able to make several supervised models that could give use the relationship of each factor to the tonnage of plastic waste generated. Using the unsupervised learning methods we could not find correlations between the factors and the response, but we were able to find groups that share common elements. Using these groups we are able to see which factors are grouping entities together, and see how much tonnage of plastic waste was generated for each group.

## Data Cleaning

### Missing Data

|       The compiled dataset had a high amount of missing values; very few observation had values for every factor. Out of 244 observations 76 observations (31.1%) were missing the response value: 'Plastic waste generation in tonnes'. A complied list of all the proportion of missing values for each variable form the complied dataset is shown in (FIGURE: table of NA's by factor). To cut out observation that had missing values would have reduced the dataset by too much to make a statistically meaningful analysis and *source* showed that data that was missing up-to $40%$ of the response variable were still able use imputation methods successfully. Thus, to preserve as much data as possible we explored several methods of imputation. Random-Forest imputation was used as well as Cart method.

```{r, echo=FALSE}

read_csv("Data Clean/NA_table.csv", show_col_types = FALSE)

```

|       There was concern about data that showed Informed missing; entities with smaller GPD were more likely to have missing data about plastic waste generation. Furthermore, politically unstable nation have a harder time to gather the appropriate waste management data. However we were not able to add appropriate metrics to our dataset to account for politician instability for each entity. Yet, since the alternative is forgoing any analysis we preceded with the imputation but remained mindful of possible short-comings of the analysis due to the possibility of informed missing data.

### Correlated Factors

|       Several of the factors that were used in this study are directly or indirectly related to each other. Examples include per-capita metrics as well as population and GDP. In order to control for all of the possible confounding interactions several models of ranging complexity were be constructed using different set of factors that showed the effects of those factors. Using this method we were able to construct a final model that used the best functionality for each factor, and added interaction terms when needed.

## Results

| 

## Future Works

|       This investigation was limited by the data that was available at the time. Information on plastic waste generated was limited to the single year of 2010 due to this limitation, losing the ability to analysis the trends of plastic waste over time. Future work on this topic should include historical data to understand possible inflection points where the trend of plastic waste generated changed, possibly revealing important events.

|       Political structures and stability were another interesting factors that influence plastic waste, but we did not have the appropriate metrics to include an investigation of these factors. Future works on this topic should add metrics for political structures and political stability to properly account their effect.

|       Geography is another important factor that could not be properly accounted for in this study. In this study the only geographical information that was counted for was land area, however this does not account for coastal or land-locked entities, nor the predominate climate of that entity. Waste management programs must adjust its practices to suite the environment which it is in. Future studies should introduce categories of different geographic feature to the dataset to study the statistical effects of geography.

## Conclution

|      This study shows...

## Apendix
