---
title: "Imputation and EDA"
author: "Ifechi Umeh"
date: "4/11/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(naniar)
library(mice)
library(miceadds)
library(ranger)
set.seed(79069084)
```


# Parsing Data 

```{r}
data <- read_csv("Data Clean/plastic_1.csv")

data <- data %>% naniar::replace_with_na(replace = list(gdp_per_capita = "no data")) # some values labelled "no data"
sum(is.na(data$gdp_per_capita))    #Number of NAs: 50; proportion missing: 0.204918

data$gdp_per_capita <- as.numeric(data$gdp_per_capita)

# data <- select(data, -c("Entity"))

names(data) <- c("entity", "waste_gen", "population", "growth_rate", "area", "density", "fert_rate", "life_exp", "mort_rate", "per_cap_waste", "misman_waste", "share_misman", "per_cap_gdp")
```


# Visualizing Missing Data 

```{r, eval=FALSE}
summary(data)

md.pattern(data) #visualizing missing values

vis_miss(data, cluster = F, sort_miss = T, show_perc = T,
  show_perc_col = F, large_data_size = 9e+05)
``` 



# Imputing Data

```{r}
imputed_data <- mice(data, 
                     m = 31,          # m selected using max(5, 100p) where p=0.3114754 (highest prop. of NAs)  
                     method = 'rf',   # uses ranger package
                     seed = 79069084) # seed selected using sample(1:999999999, size = 1)
```
Error from using "pmm" method for imputation: "Error in solve.default(xtx + diag(pen)) : system is computationally singular: reciprocal condition number = 6.10832e-20"

```{r, eval=FALSE}
table_10 <- complete(imputed_data, 10)    # sample(1:31, size=1)
```


# Randomly selected table for visualizations

```{r}
# randomly selected table for visualizations
ggplot(data=table_10, aes(x=log(population), y=log(waste_gen))) +
  geom_point()
```

```{r}
ggplot(data=table_10, aes(x=log(area), y=log(waste_gen))) +
  geom_point()
```

```{r}
ggplot(data=table_10, aes(x=log(misman_waste), y=log(waste_gen))) +
  geom_point()
```

```{r}
ggplot(data=table_10, aes(x=log(per_cap_waste), y=log(waste_gen))) +
  geom_point()
```

```{r}
ggplot(data=table_10, aes(x=log(share_misman), y=log(waste_gen))) +
  geom_point()
```

```{r}
ggplot(data=table_10, aes(x=log(per_cap_gdp), y=log(waste_gen))) +
  geom_point()
```


Population, area, misman_waste, per_cap_waste, share_misman

# Correlation

```{r, eval=FALSE}
# correlation matrix
correlations <- miceadds::micombine.cor(mi.res=imputed_data, variables = c(2:12))

old_school<-format(correlations$p, scientific=FALSE)
correlations["p_value"] <- NA; correlations$p_value <- old_school; 
correlations
```

population: area and misman_waste

# Stepwise Selection

```{r}
scope <- list(upper = ~ growth_rate + area + density + fert_rate + life_exp + mort_rate + per_cap_waste + misman_waste + share_misman + per_cap_gdp,
              lower = ~1)
expr <- expression(f1 <- lm(waste_gen ~ 1),
                   f2 <- step(f1, scope = scope, trace=0))
fit <- with(imputed_data, expr)
```

```{r}
formulas <- lapply(fit$analyses, formula)
terms <- lapply(formulas, terms)
votes <- unlist(lapply(terms, labels))
table(votes)
```

# Random Forest
```{r}
with(data = imputed_data, expr = randomForest(waste_gen~.,data=table, mtry=4, importance=TRUE,ntree=500))

```


# Fitting models

```{r}
fit_1 <- with(data = imputed_data, expr = lm(waste_gen ~ area + misman_waste + per_cap_waste + share_misman))

fit_combine <- pool(fit_1)
summary(fit_combine)
```

```{r}
fit_without <- with(imputed_data, lm(waste_gen ~ area + misman_waste + per_cap_waste ))
fit_with <- with(imputed_data, lm(waste_gen ~ area + misman_waste + per_cap_waste + share_misman))
D1(fit_with, fit_without)
```


```{r}
fit_without2 <- with(imputed_data, lm(waste_gen ~ area + misman_waste))
fit_with2 <- with(imputed_data, lm(waste_gen ~ area + misman_waste + per_cap_waste))
D1(fit_with2, fit_without2)
```


```{r}
fit_2 <- with(data = imputed_data, expr = lm(waste_gen ~ area + misman_waste + per_cap_waste))

fit2_combine <- pool(fit_2)
summary(fit2_combine)
```











