---
title: "Gathering Data"
author: "Ifechi Umeh"
date: "3/28/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
```


*states vs countries. We need to define what a country is,
*use entity instead of country

```{r, eval=FALSE}
waste_total <- read_csv("Data Raw/plastic-waste-generation-total.csv")

waste_per_cap <- read_csv("Data Raw/plastic-waste-per-capita.csv")
waste_per_cap <-  waste_per_cap %>% 
                    select(c("Entity", "Per capita plastic waste (kg/person/day)"))

waste_totgbl_mismanage <- read_csv("Data Raw/mismanaged-waste-global-total.csv")
waste_totgbl_mismanage <-  waste_totgbl_mismanage %>% 
                             select(c("Entity", "Mismanaged waste (% global total)"))

waste_inadequate_manage <- read_csv("Data Raw/inadequately-managed-plastic.csv")
waste_inadequate_manage <-  waste_inadequate_manage %>% 
                            select(c("Entity", "Share of plastic inadequately managed"))
```

*Plastic data from Our World in Data

<<<<<<< HEAD
```{r, eval=FALSE}
=======
```{r}
>>>>>>> 31b5a6386a4662ef9c7162cc51000080db1a80b6
population_ <- read_csv("Data Raw/export.csv")
pop <- read_csv("Data Raw/export.csv")
pop <- pop %>% 
        select(c("Country/Area Name", "Population", "Annual Growth Rate %", "Area (sq km)", "Density (per sq km)", "Total Fertility Rate", "Life Expectancy at Birth", "Under-5 Mortality Rate"))
```

Note: A lot of the N/a's in the wiki webpage are smaller Islands nations/ territories. I would be a little uncomfortable with imputing this data since it is Informed Missing. I found that we could find some of those missing values

*changed the population since Wikipedia is not a reliable source. Data from Census Bureau

bahamas

bahamas, the

```{r, eval=FALSE}
country_gdp <- read_csv("Data Raw/Country-GPD-per-capita.csv")

country_gdp_capita <- country_gdp %>% transmute(Entity = `GDP per capita, current prices
 (U.S. dollars per capita)`,
 gdp_per_capita = `2010`)

gdp_per_cap <- country_gdp_capita[c(2:197),] #to ensure we have just the countries and not regions
```

*"GDP per capita, current prices (U.S. dollars per capita)": Data from IMF

"Gross domestic product is the most commonly used single measure of a country's overall economic activity. It represents the total value at current prices of final goods and services produced within a country during a specified time period divided by the average population for the same one year.

Source: World Economic Outlook (October 2021)" https://www.imf.org/external/datamapper/NGDPDPC@WEO/OEMDC/ADVEC/WEOWORLD 


<<<<<<< HEAD
```{r, eval=FALSE}
=======
```{r}
>>>>>>> 31b5a6386a4662ef9c7162cc51000080db1a80b6
tab1 <- full_join(waste_total, pop, by = c("Entity" = "Country/Area Name"))
tab2 <- full_join(tab1, waste_per_cap, by = c("Entity"))
tab3 <- full_join(tab2, waste_totgbl_mismanage, by = c("Entity"))
tab4 <- full_join(tab3, waste_inadequate_manage, by = c("Entity"))
tab5 <- full_join(tab4, gdp_per_cap, by = c("Entity"))
#tab6 <- left_join(tab5, country_gdp_capita, by = "Entity") not sure what year is the USD in.
```

```{r, eval=FALSE}
#write_csv(tab5, "Data Clean/master_table.csv")
#write_csv(tab5, "Data Clean/data_1.csv")
```

*need to fix: some countries are spelt differently in different sources leading to duplicate rows that have NAs.
*very difficult to find data on total plastic produced by country in 2010.



```{r, eval=FALSE}
tab6 <- tab5
# Tidying up Entity names
tab6$Entity <- str_replace(tab6$Entity, ".*Bahamas.*", "Bahamas")
tab6$Entity <- str_replace(tab6$Entity, ".*Brunei.*", "Brunei")
tab6$Entity <- str_replace(tab6$Entity, "China, People's Republic of", "China")
tab6$Entity <- str_replace(tab6$Entity, "^Congo$", "Republic of Congo")
tab6$Entity <- str_replace(tab6$Entity, "Congo \\(Brazzaville\\)", "Republic of Congo")
tab6$Entity <- str_replace(tab6$Entity, "Congo, Republic of", "Republic of Congo")
tab6$Entity <- str_replace(tab6$Entity, "Congo \\(Kinshasa\\)", "Democratic Republic of Congo")
tab6$Entity <- str_replace(tab6$Entity, "Congo, Dem. Rep. of the", "Democratic Republic of Congo")
tab6$Entity <- str_replace(tab6$Entity, ".*C[ôo]te.*", "Cote d'Ivoire")
tab6$Entity <- str_replace(tab6$Entity, ".*Cura[cç]ao.*", "Curacao")
tab6$Entity <- str_replace(tab6$Entity, ".*Fa.?roe.*", "Faeroe Islands")
tab6$Entity <- str_replace(tab6$Entity, ".*Gambia.*", "The Gambia")
tab6$Entity <- str_replace(tab6$Entity, ".*Hong Kong.*", "Hong Kong")
tab6$Entity <- str_replace(tab6$Entity, "Korea, North", "North Korea")
tab6$Entity <- str_replace(tab6$Entity, "Korea, South", "South Korea")
tab6$Entity <- str_replace(tab6$Entity, "Korea, Republic of", "South Korea")
tab6$Entity <- str_replace(tab6$Entity, ".*Kyrgyz.*", "Kyrgyzstan")
tab6$Entity <- str_replace(tab6$Entity, ".*Lao.*", "Laos")
tab6$Entity <- str_replace(tab6$Entity, ".*Maca.*", "Macao")
tab6$Entity <- str_replace(tab6$Entity, ".*Micronesia.*", "Micronesia")
tab6$Entity <- str_replace(tab6$Entity, ".*Russia.*", "Russia")
tab6$Entity <- str_replace(tab6$Entity, ".*Saint Helena.*", "Saint Helena")
tab6$Entity <- str_replace(tab6$Entity, "São Tomé and Príncipe", "Sao Tome and Principe")
tab6$Entity <- str_replace(tab6$Entity, ".*Sint Maarten.*", "Sint Maarten")
tab6$Entity <- str_replace(tab6$Entity, ".*Slovak.*", "Slovakia")
tab6$Entity <- str_replace(tab6$Entity, ".*South Sudan.*", "South Sudan")  # South Sudan may need to be removed
tab6$Entity <- str_replace(tab6$Entity, ".*Taiwan.*", "Taiwan")
tab6$Entity <- str_replace(tab6$Entity, "Virgin Islands, British", "British Virgin Islands")


tab7 <- tab6 %>% 
          group_by(Entity) %>% 
          fill(everything(), .direction = "downup") %>% 
          slice(n=1)
```

* Congo Brazzaville (Republic of Congo) and Congo Kinshasa (Democratic Republic of Congo)
* country code: COG -> Republic of Congo; COD -> Democratic Republic of Congo
*Note: South Sudan and Sudan were one country until they split in 2011. Issue because data is 2010. Possibly take out the South Sudan row
*Not sure what to do about west bank and west bank and gaza


```{r, eval=FALSE}
# NA EDA: Number of observations: 244
sum(is.na(tab7$`Plastic waste generation (tonnes, total)`)) #Number of NAs: 76; proportion missing: 0.3114754
sum(is.na(tab7$Population))                                 #Number of NAs: 17; proportion missing: 0.06967213
sum(is.na(tab7$`Annual Growth Rate %`))                     #Number of NAs: 18; proportion missing: 0.07377049 #US missing 
sum(is.na(tab7$`Area (sq km)`))                             #Number of NAs: 17; proportion missing: 0.06967213
sum(is.na(tab7$`Density (per sq km)`))                      #Number of NAs: 17; proportion missing: 0.06967213
sum(is.na(tab7$`Total Fertility Rate`))                     #Number of NAs: 18; proportion missing: 0.07377049
sum(is.na(tab7$`Life Expectancy at Birth`))                 #Number of NAs: 18; proportion missing: 0.07377049
sum(is.na(tab7$`Under-5 Mortality Rate`))                   #Number of NAs: 18; proportion missing: 0.07377049
sum(is.na(tab7$`Per capita plastic waste (kg/person/day)`)) #Number of NAs: 58; proportion missing: 0.2377049
sum(is.na(tab7$`Mismanaged waste (% global total)`))        #Number of NAs: 58; proportion missing: 0.2377049
sum(is.na(tab7$`Share of plastic inadequately managed`))    #Number of NAs: 58; proportion missing: 0.2377049
sum(is.na(tab7$gdp_per_capita))                             #Number of NAs: 48; proportion missing: 0.1967213
```

<<<<<<< HEAD
```{r, eval=FALSE}
=======
```{r}
>>>>>>> 31b5a6386a4662ef9c7162cc51000080db1a80b6
# Remove Code and Year
tab8 <- select(tab7, -c("Code", "Year"))
```


<<<<<<< HEAD
```{r, eval=FALSE}
=======
```{r}
>>>>>>> 31b5a6386a4662ef9c7162cc51000080db1a80b6
write_csv(tab8, "Data Clean/plastic_1.csv")
```
